Response to reviewers

### 2025-07-04 src/docs_paper-jtb/paper.Rmd
Figure update text
Move ADREPORT into main text and update reply

(ref:ks-summary) The average Kolmogorov-Smirnov (KS) test statistic for each latent field parameter of the Naomi model. Vectors of parameters were grouped together. For points above the dashed line at zero, performance of GEB was better. For points below the dashed line, performance of GPCA-AGHQ was better. Most notably, for the latent field parameters $\texttt{ui\_lambda\_x}$ the average test statistic for GEB was substantially higher than for GPCA-AGHQ. This length 32 parameter $\mathbf{u}_x^{\lambda}$ plays a key role in the ART attendance component of the Naomi (Section \@ref(sec:art)).

@@

Comparison to further Bayesian inference methods could prove beneficial.
Four possibilities stand out as being particularly valuable.
First, there exist other quadrature rules for moderate dimension, such as the Box-Wilson central composite design [@box1992experimental].
It would be of interest to compare INLA with a PCA-AGHQ rule to INLA with other such quadrature rules.
Second, rather than use quadrature to integrate the marginal Laplace approximation, an alternative approach is to run HMC [@monnahan2018no; @margossian2020hamiltonian].
This method can be run by using `tmbstan` and setting `laplace = TRUE`.
When run to convergence, inferential error of this method would solely be due to the Laplace approximation, helping to clarify the extent to which the inferential error of INLA is attributable to the quadrature grid.
Third, NUTS is not especially well suited to sampling from Gaussian latent field models like Naomi.
Other MCMC algorithms, such as blocked Gibbs sampling [@geman1984stochastic] or slice sampling [@neal2003slice], could be considered.
It may be difficult to implement such algorithms using `TMB`.
Many MCMC algorithms are implemented and customisable (including, for example, the choice of block structure) within the `NIMBLE` probabilistic programming language (PPL) [@de2017programming].
Additionally, direct implementation of Naomi within the Stan PPL could alter the speed by as much as plus or minus 50% [@monnahan2018no] and would give access to other algorithms implemented in Stan such as variational inference [@kucukelbir2015automatic] and Pathfinder [@zhang2022pathfinder].
However, requiring rewriting the Naomi model log-posterior outside of `TMB` would be a subtantial downside.
Finally, it would be of substantial interest to implement the Naomi model using the iterative INLA method via `inlabru`.
However, as `inlabru`, like `R-INLA`, is based on a formula interface, it may not be possible to do so directly.

@@

PCA-AGHQ is a sensible approach to allocating more computational effort to dimensions which contribute more to the integral in question.
Its application to Naomi surfaced several ideas for improvement and challenges to be overcome.
First, the amount of variation explained in the Hessian matrix may not be of direct interest.
For the Naomi model, interest is in the effect of including each dimension on the relevant model outputs.
As such, using alternative measures of importance from sensitivity analysis, such as Shapley values [@shapley1953value] or Sobol indices, could be preferable.
Second, use of PCA is challenging when the dimensions have different scales.
For the Naomi model, logit-scale hyperparameters were systematically favoured over those on the log-scale.
Third, when the quadrature rule is used within an INLA algorithm, it is more important to allocate quadrature nodes to those hyperparameter marginals which are non-Gaussian.
This is because the Laplace approximation is exact when the integrand is Gaussian, so a single quadrature node is sufficient.
The difficulty is knowing in advance which marginals will be non-Gaussian.
This could be done if there were a cheap way to obtain posterior means, which could then be compared to posterior modes obtained using optimisation.
Another approach would be to measure the fit of marginal samples from a cheap approximation, like EB.
The measures of fit would have to be for marginals, ruling out approaches like PSIS [@yao2018yes] which operate on joint distributions.
Finally, it is likely possible to achieve better performance by pruning (dropping points where the weight is below some threshold) or prerotation (before adaption, in order to e.g. achieve better coverage along the principal axis) of the quadrature grid [@jackel2005note].

>>>>

(ref:ks-summary) The average Kolmogorov-Smirnov (KS) test statistic for each latent field parameter of the Naomi model. Vectors of parameters were grouped together. For points above the dashed line at zero, performance of GEB was better. For points below the dashed line, performance of GPCA-AGHQ was better. Most notably, for the latent field parameters $\texttt{ui\_lambda\_x}$ the average test statistic for GEB was substantially higher than for GPCA-AGHQ. This length 32 parameter $\mathbf{u}_x^{\lambda}$ plays a key role in modelling ART attendance (Section \@ref(sec:art)).

@@

Comparison to further Bayesian inference methods could prove beneficial.
Four possibilities stand out as being particularly valuable.
First, there exist other quadrature rules for moderate dimension, such as the Box-Wilson central composite design [@box1992experimental].
It would be of interest to compare INLA with a PCA-AGHQ rule to INLA with other such quadrature rules.
Second, rather than use quadrature to integrate the marginal Laplace approximation, an alternative approach is to run HMC [@monnahan2018no; @margossian2020hamiltonian].
This method can be run by using `tmbstan` and setting `laplace = TRUE`.
When run to convergence, inferential error of this method would solely be due to the Laplace approximation, helping to clarify the extent to which the inferential error of INLA is attributable to the quadrature grid.
Third, NUTS is not especially well suited to sampling from Gaussian latent field models like Naomi.
Other MCMC algorithms, such as blocked Gibbs sampling [@geman1984stochastic] or slice sampling [@neal2003slice], could be considered.
It may be difficult to implement such algorithms using `TMB`.
Many MCMC algorithms are implemented and customisable (including, for example, the choice of block structure) within the `NIMBLE` probabilistic programming language (PPL) [@de2017programming].
Additionally, direct implementation of Naomi within the Stan PPL could alter the speed by as much as plus or minus 50% [@monnahan2018no] and would give access to other algorithms implemented in Stan such as variational inference [@kucukelbir2015automatic] and Pathfinder [@zhang2022pathfinder].
However, requiring rewriting the Naomi model log-posterior outside of `TMB` would be a substantial downside.
Finally, it would be of substantial interest to implement the Naomi model using the iterative INLA method via `inlabru`.
However, as `inlabru`, like `R-INLA`, is based on a formula interface, it may not be possible to do so directly.

@@

PCA-AGHQ is a sensible approach to allocating more computational effort to dimensions which contribute more to the integral in question.
Its application to Naomi surfaced several ideas for improvement and challenges to be overcome.
First, the amount of variation explained in the Hessian matrix may not be of direct interest.
For the Naomi model, interest is in the effect of including each dimension on the relevant model outputs.
As such, using alternative measures of importance from sensitivity analysis, such as Shapley values [@shapley1953value] or Sobol indices, could be preferable.
This may be achievable by applying the `ADREPORT` function to model outputs within the `TMB` template to obtain gradients.
Second, use of PCA is challenging when the dimensions have different scales.
For the Naomi model, logit-scale hyperparameters were systematically favoured over those on the log-scale.
Third, when the quadrature rule is used within an INLA algorithm, it is more important to allocate quadrature nodes to those hyperparameter marginals which are non-Gaussian.
This is because the Laplace approximation is exact when the integrand is Gaussian, so a single quadrature node is sufficient.
The difficulty is knowing in advance which marginals will be non-Gaussian.
This could be done if there were a cheap way to obtain posterior means, which could then be compared to posterior modes obtained using optimisation.
Another approach would be to measure the fit of marginal samples from a cheap approximation, like EB.
The measures of fit would have to be for marginals, ruling out approaches like PSIS [@yao2018yes] which operate on joint distributions.
Finally, it is likely possible to achieve better performance by pruning (dropping points where the weight is below some threshold) or prerotation (before adaption, in order to e.g. achieve better coverage along the principal axis) of the quadrature grid [@jackel2005note].

### 2025-07-04 src/docs_paper-jtb/revisions.Rmd
Move ADREPORT into main text and update reply

It is possible to obtain derivative information about a quantity in a `TMB` model using `ADREPORT`.

>>>>

We agree this is an interesting area for future work, and have added a sentence to Section 6.1.2. discussing the possibility of doing this in `TMB`:

> This may be achievable by applying the `ADREPORT` function to model outputs within the `TMB` template to obtain gradients.
